\documentclass[14pt]{extarticle}

\usepackage{sourceserifpro}
\usepackage[margin=0.75in]{geometry}
\usepackage{microtype}
\usepackage{enumitem}


\pagenumbering{gobble}


\begin{document}

\begin{center}
{\large\scshape\bfseries Eight Stages of White Settler-Colonial Denial}
\end{center}
\vspace{1em}

\begin{enumerate}[leftmargin=0in]
\item {\bfseries``They didn't exist''\hfill(terra nullius)}

Complete denial of Indigenous presence in a given area (country, province, etc).
Includes denial of Indigeneity, e.g. ``Indigenous Peoples are Settlers too''.

\item {\bfseries``If they did, they weren't here''\hfill(terra nullius)}

Denial that Indigenous People inhabit/travel/harvest/exist in a specific area.
Often based on euro-centric definitions of evidence of occupation.

\item {\bfseries``If they were, they didn't use the land''\hfill(doctrine of discovery)}

Denial that Indigenous People have a connection with the Land.
Often based on euro-centric worldviews of the land as something to be owned and extracted.

\item {\bfseries``If they did, they didn't deserve it''\hfill(great chain of being)}

Denial that Indigenous People have rights to their Lands.
Often based on euro-centric value judgements of ``primitive vs. civilized'', ``nomadic vs. sedentary''.

\item {\bfseries``If they did, they lost it''\hfill(right of conquest)}

Denial that Indigenous People retain their rights to their Lands.
Often based on colonially imposed European systems of law / ``might makes right'' worldviews.

\item {\bfseries``If they didn't, it doesn't matter anymore''\hfill(Westphalian sovereignty)}

Denial that Indigenous Rights are still binding and take precedence.
Often based on false claims of supremacy of colonial legal institutions and systems.

\item {\bfseries``If it does, we need to move on''.\hfill(liberalism)}

Denial that violations of Indigenous Rights requires redress.
Often based on claims redress is ``disruptive/unfair/reverse racism'' \& false calls for ``equality''.

\item {\bfseries``If we can't, we are you''\hfill(self-indigenization)}

Denial of separateness of Indigenous Peoples and Rights.
Often based on attempts to reduce Indigenous Rights to Human Rights, claim Indigeneity, etc.
\end{enumerate}

\vfill



{\footnotesize CC SA 4.0\hfill``Eight Stages of Settler Denial''. Revision I}

\end{document}
